Visualisations for blog on SCADR website

### 2021-04-16 visualisations_for_blog.Rmd
modified blog & paper scripts to load the workspace with visualsiations included
Added age group to blog visualisations

```{r}
load(file = "./workspace.RData")
```

@@

@@

>>>>

```{r}
load(file = "./workspace_with_visualisations.RData")
```

@@

### Table showing increases by area

@@

## Home deaths by age group

```{r, fig.width=12, fig.height=9, warning=FALSE}
removed_under_15_text <- glue::glue("Excluding individuals aged <15 at death, total N={merged_deaths_age %>% filter(age %in% c('0','1-14') & place_of_death == 'Home & other non-institution') %>% .$deaths_all_causes %>% sum} for weeks 1-{max(merged_deaths_age %>% filter(year==2020) %>% .$week_number)} in 2020 and 1-{max(merged_deaths_age %>% filter(year==2021) %>% .$week_number)} in 2021.")

## to put the arrows explaining historic range etc. only in one panel, need to construct a tibble to pass as data
annotations_figure_home_deaths_by_age <- tribble(
   ~geom, ~place_of_death, ~age, ~label, ~x, ~y, ~xend, ~yend,
  "text", "Home & other non-institution", "45-64",  "Historic range (min & max)", 17, 20, NA, NA,
  "curve", "Home & other non-institution", "45-64", NA, 22, 25, 25, 43,
  "text", "Home & other non-institution", "45-64", "Historic mean", 42, 20, NA, NA,
  "curve", "Home & other non-institution", "45-64", NA, 42, 25, 44, 64
) %>% 
  mutate(across(.cols = c("x","xend"), .fns = ~compute_start_date_from_week_number(week_number = .x, year_number = 2020)))

(figure_deaths_at_home_by_age <-
  merged_deaths_age %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  filter(!age %in% c('0','1-14')) %>%
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  # ggplot(aes(x = week_number_run_over)) +
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", size = 0.2) +
  # geom_vline(xintercept = 54, linetype = "dotted") +
  annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) + 
  # facet_wrap(~age, scales = "free_y") +
  facet_wrap(~age) +
  # theme_minimal() +

  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
   theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = "Week number",
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(removed_under_15_text, source_nrs_text), collapse="\n"),
       colour = "Cause of death"
       ) +
  ## annotate historic range
  geom_text(data = annotations_figure_home_deaths_by_age %>% filter(geom=="text"), aes(x=x, y=y, label=label),  colour = "grey20", size = 8 / (14/5)) +  # the size is tricky
  geom_curve(data = annotations_figure_home_deaths_by_age %>% filter(geom=="curve"), aes(x=x, y=y, xend=xend, yend=yend), colour = c("#4477aa", "grey20"), alpha = c(0.5, 1), linetype = c("solid","dashed"), curvature = 0.20, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=10, label = "2021", hjust = 0, colour = "grey40") +
  NULL
)

save_output_file(filename = paste0(dir_blog,"/figure_deaths_at_home_by_age"), extensions = c(".pdf",".svg",".png"), plot = figure_deaths_at_home_by_age, width = 12, height = 9, dpi = 300, units = "in")
```

### Table showing increases by age

```{r}
table_increase_home_deaths_by_age <-
  merged_deaths_age %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  group_by(year, place_of_death, age, week_number) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
  left_join(
    past_deaths_age %>% filter(year != "average") %>% group_by(week_number, place_of_death, age) %>% summarise(
                                       average_weekly_deaths = sum(number_of_deaths) / 5,
                                       .groups = "drop"
                                     ),
    by = c("week_number", "place_of_death", "age")
  ) %>% 
  group_by(place_of_death, age) %>%
  summarise(
    period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
    deaths_all_causes = sum(deaths_all_causes),
    deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
    .groups = "drop"
  ) %>%
  mutate(percent_increase_deaths_over_historic = (deaths_all_causes / mean_annual_deaths_2015_2019) - 1) %>%
  arrange(percent_increase_deaths_over_historic) %>%
  mutate(
    `Percent increase` = scales::percent(percent_increase_deaths_over_historic, accuracy = 0.1)
    ) %>%
  filter(!age %in% c("0", "1-14")) %>%  # remove the v.small groups
  select(`Age Group` = age, `Home deaths` = deaths_all_causes, `Percent increase`)

write_csv(x = table_increase_home_deaths_by_age, file = paste0(dir_blog,"/table_increase_home_deaths_by_age.csv"))
```
